---
title: "Exercise 5"
format: html
editor: visual
---

```{r}
#load the packages needed for the tutorial
pacman::p_load(tidyverse, stringr, tidytext, topicmodels, gutenbergr, scales, tm, ggthemes, readr, quanteda, quanteda.textmodels, matrix, slam, tm)
```

```{r}
#We'll be using Alexis de Tocqueville's "Democracy in America." We will download these data both vol 1 and 2 and combine them into one data frame. For this we'll be using the gutenbergr package which allows the user to download text data from 60,000 out of copyright books. The ID for each book appears in the url for the book selected after a search on gutenberg.org/ebooks
devtools::install_github("matthewjdenny/preText")
library(preText)
```

```{r}
#Download the versions that we need
tocq <- gutenberg_download(c(815, 816),
                           meta_fields = "author")

#Once we have read in these data, we convert it into a different data shape: the document-term-matrix. We also create a new columns, which we call “booknumber” that recordss whether the term in question is from Volume 1 or Volume 2. To convert from tidy into “DocumentTermMatrix” format we can first use unnest_tokens() as we have done in past exercises, remove stop words, and then use the cast_dtm() function to convert into a “DocumentTermMatrix” object.

tocq_words <- tocq %>%
  mutate(booknumber = ifelse(gutenberg_id==815, "DiA1", "DiA2")) %>%
  unnest_tokens(word, text) %>%
  filter(!is.na(word)) %>%
  count(booknumber, word, sort = TRUE) %>%
  ungroup() %>%
  anti_join(stop_words)

tocq_dtm <- tocq_words %>%
  cast_dtm(booknumber, word, n)

tm::inspect(tocq_dtm)
#We see here that the data are now stored as a “DocumentTermMatrix.” In this format, the matrix records the term (as equivalent of a column) and the document (as equivalent of row), and the number of times the term appears in the given document. Many terms will not appear in the document, meaning that the matrix will be stored as “sparse,” meaning there will be a preponderance of zeroes. Here, since we are looking only at two documents that both come from a single volume set, the sparsity is relatively low (only 27%). In most applications, the sparsity will be a lot higher, approaching 99% or more.

#Estimating our topic model is then relatively simple. All we need to do if specify how many topics that we want to search for, and we can also set our seed, which is needed to reproduce the same results each time (as the model is a generative probabilistic one, meaning different random iterations will produce different results).

tocq_lda <- LDA(tocq_dtm, k = 10, control = list(seed = 1234))

tocq_topics <-tidy(tocq_lda, matrix= "beta")

head(tocq_topics, n=10)

```

```{r}
#We now have data stored as one topic-per-term-per-row. The betas listed here represent the probability that the given term belongs to a given topic. So, here, we see that the term “democratic” is most likely to belong to topic 4. Strictly, this probability represents the probability that the term is generated from the topic in question.

#We can then plots the top terms, in terms of beta, for each topic as follows:

tocq_top_terms <- tocq_topics %>%
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)

tocq_top_terms %>%
  mutate(term = reorder_within(term, beta, topic)) %>%
  ggplot(aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free", ncol = 4) +
  scale_y_reordered() +
  theme_tufte(base_family = "Helvetica")

#But how do we actually evaluate these topics? Here the topics seem all pretty similar


```

```{r}
#Evaluating the topic model
#Well, one way to evaluate the performance of unspervised forms of classification is by testing our model on an outcome that is already known.

#Here, two topics that are most obvious are the ‘topics’ Volume 1 and Volume 2 of Tocqueville’s “Democracy in America.” Volume 1 of Tocqueville’s work deals more obviously with abstract constitutional ideas and questions of race; Volume 2 focuses on more esoteric aspects of American society. Listen an “In Our Time” episode with Melvyn Bragg discussing Democracy in America here.

#Given these differences in focus, we might think that a generative model could accurately assign to topic (i.e., Volume) with some accuracy.

#21.3.1 Plot relative word frequencies
#First let’s have a look and see whether there really are words obviously distinguishing the two Volumes.

tidy_tocq <- tocq %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words)

#count most common words in both

tidy_tocq %>%
  count(word, sort= TRUE)

bookfreq <- tidy_tocq %>%
  mutate(booknumber = ifelse(gutenberg_id==815, "DiA1", "DiA2")) %>%
  mutate(word = str_extract(word, "[a-z']+")) %>%
  count(booknumber, word) %>%
  group_by(booknumber) %>%
  mutate(proportion = n / sum(n)) %>% 
  select(-n) %>% 
  spread(booknumber, proportion)

ggplot(bookfreq, aes(x = DiA1, y = DiA2, color = abs(DiA1 - DiA2))) +
  geom_abline(color = "gray40", lty = 2) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient(limits = c(0, 0.001), low = "darkslategray4", high = "gray75") +
  theme_tufte(base_family = "Helvetica") +
  theme(legend.position="none", 
        strip.background = element_blank(), 
        strip.text.x = element_blank()) +
  labs(x = "Tocqueville DiA 2", y = "Tocqueville DiA 1") +
  coord_equal()

```

We see that there do seem to be some marked distinguishing characteristics. In the plot above, for example, we see that more abstract notions of state systems appear with greater frequency in Volume 1 while Volume 2 seems to contain words specific to America (e.g., “north” and “south”) with greater frequency. The way to read the above plot is that words positioned further away from the diagonal line appear with greater frequency in one volume versus the other.

### **21.3.2 Split into chapter documents**

In the below, we first separate the volumes into chapters, then we repeat the same procedure as above. The only difference now is that instead of two documents representing the two full volumes of Tocqueville’s work, we now have 132 documents, each representing an individual chapter. Notice now that the sparsity is much increased: around 96%.

```{r}
tocq <- tocq %>%
  filter(!is.na(text))

# Divide into documents, each representing one chapter
tocq_chapter <- tocq %>%
  mutate(booknumber = ifelse(gutenberg_id==815, "DiA1", "DiA2")) %>%
  group_by(booknumber) %>%
  mutate(chapter = cumsum(str_detect(text, regex("^chapter ", ignore_case = TRUE)))) %>%
  ungroup() %>%
  filter(chapter > 0) %>%
  unite(document, booknumber, chapter)

# Split into words
tocq_chapter_word <- tocq_chapter %>%
  unnest_tokens(word, text)

# Find document-word counts
tocq_word_counts <- tocq_chapter_word %>%
  anti_join(stop_words) %>%
  count(document, word, sort = TRUE) %>%
  ungroup()

tocq_word_counts
```

```{r}
#Cast into DTM format for LDA analysis
tocq_chapters_dtm <- tocq_word_counts %>%
  cast_dtm(document, word, n)
tm::inspect(tocq_chapters_dtm)
```

We then re-estimate the topic model with this new DocumentTermMatrix object, specifying k equal to 2. This will enable us to evaluate whether a topic model is able to generatively assign to volume with accuracy.

```{r}
tocq_chapters_lda <-LDA(tocq_chapters_dtm, k = 2, control = list(seed= 1234))
```

After this, it is worth looking at another output of the latent dirichlet allocation procedure. The γ probability represents the per-document-per-topic probability or, in other words, the probability that a given document (here: chapter) belongs to a particular topic (and here, we are assuming these topics represent volumes).

The gamma values are therefore the estimated proportion of words within a given chapter allocated to a given volume.

```{r}
tocq_chapters_gamma <- tidy(tocq_chapters_lda, matrix = "gamma")
tocq_chapters_gamma
```

### **21.3.3 Examine consensus**

Now that we have these topic probabilities, we can see how well our unsupervised learning did at distinguishing the two volumes generatively just from the words contained in each chapter.

```{r}
# First separate the document name into title and chapter

tocq_chapters_gamma <- tocq_chapters_gamma %>%
  separate(document, c("title", "chapter"), sep = "_", convert = TRUE)

tocq_chapter_classifications <- tocq_chapters_gamma %>%
  group_by(title, chapter) %>%
  top_n(1, gamma) %>%
  ungroup()

tocq_book_topics <- tocq_chapter_classifications %>%
  count(title, topic) %>%
  group_by(title) %>%
  top_n(1, n) %>%
  ungroup() %>%
  transmute(consensus = title, topic)

tocq_chapter_classifications %>%
  inner_join(tocq_book_topics, by = "topic") %>%
  filter(title != consensus)
```

```{r}
# Look document-word pairs were to see which words in each documents were assigned
# to a given topic

assignments <- augment(tocq_chapters_lda, data = tocq_chapters_dtm)
assignments
```

```{r}
assignments <- assignments %>%
  separate(document, c("title", "chapter"), sep = "_", convert = TRUE) %>%
  inner_join(tocq_book_topics, by = c(".topic" = "topic"))

assignments %>%
  count(title, consensus, wt = count) %>%
  group_by(title) %>%
  mutate(percent = n / sum(n)) %>%
  ggplot(aes(consensus, title, fill = percent)) +
  geom_tile() +
  scale_fill_gradient2(high = "red", label = percent_format()) +
  geom_text(aes(x = consensus, y = title, label = scales::percent(percent))) +
  theme_tufte(base_family = "Helvetica") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        panel.grid = element_blank()) +
  labs(x = "Book words assigned to",
       y = "Book words came from",
       fill = "% of assignments")
```

The model estimated with accuracy 77% of the chapters in volume 2 and 94% of the chapters in Volume 1

## **21.4 Validation**

In the articles by Ying et al. ([2021](https://marionlieutaud.github.io/CTA-ED/references-2.html#ref-ying_topics_2021)) and Denny and Spirling ([2018](https://marionlieutaud.github.io/CTA-ED/references-2.html#ref-denny_text_2018)) from this and previous weeks, we read about potential validation techniques.

In this section, we’ll be using the `preText` package mentioned in Denny and Spirling ([2018](https://marionlieutaud.github.io/CTA-ED/references-2.html#ref-denny_text_2018)) to see the impact of different pre-processing choices on our text. Here, I am adapting from a [tutorial](http://www.mjdenny.com/getting_started_with_preText.html) by Matthew Denny.

First we need to reformat our text into a `quanteda` corpus object.

```{r}
# load in corpus of Tocequeville text data.
corp <- corpus(tocq, text_field = "text")
# use first 10 documents for example
documents <- corp[sample(1:30000,1000)]
# take a look at the document names
print(names(documents[1:10]))
```

And now we are ready to preprocess in different ways. Here, we are including n-grams so we are preprocessing the text in 128 different ways. This takes about ten minutes to run on a machine with 8GB RAM.

```{r}
preprocessed_documents <- factorial_preprocessing(
    documents,
    use_ngrams = TRUE,
    infrequent_term_threshold = 0.2,
    verbose = FALSE)
```

We can then get the results of our pre-processing, comparing the distance between documents that have been processed in different ways.

```{r}
preText_results <- preText(
    preprocessed_documents,
    dataset_name = "Tocqueville text",
    distance_method = "cosine",
    num_comparisons = 20,
    verbose = FALSE)
```

```{r}
preText_results <- preText(
    preprocessed_documents,
    dataset_name = "Tocqueville text",
    distance_method = "cosine",
    num_comparisons = 20,
    verbose = FALSE)

#and we can plot these accordingly
preText_score_plot(preText_results)
```

21.5 Exercises

1.  Choose another book or set of books from Project Guttenberg
2.  Run your own topic model on these books, changing the k of topics, and evaluating accuracy
3.  Validate different pre processing techniques using preText on the new books of your choice

```{r}
# I am going to choose Alices's Adventures in Wonderland and Through the Looking Glass as it's one of my childhood favorite books.

#Download the versions that we need
tocq_alice <- gutenberg_download(c(11, 12),
                           meta_fields = "author")

#read in the data and then convert it to a document term matric. 
# To convert from tidy into “DocumentTermMatrix” format we can first use unnest_tokens() and then use the cast_dtm() function to convert into a “DocumentTermMatrix” object.
#Alice_1 is "Alice's Adventures in Wonderland and Alice_2 is "Through the Looking Glass"

tocq_words_alice <- tocq_alice %>%
  mutate(booknumber = ifelse(gutenberg_id==11, "Alice_1", "Alice_2")) %>%         
  unnest_tokens(word, text) %>%            
  filter(!is.na(word)) %>%
  count(booknumber, word, sort = TRUE) %>%
  ungroup() %>%
  anti_join(stop_words)
  

#view the tocq_words_alice
print(tocq_words_alice)
tocq_dtm_alice <- tocq_words_alice %>%
  cast_dtm(booknumber, word, n)

tm::inspect(tocq_dtm_alice)


# (pasting some of the explanation from above here to have it in front of me) We see here that the data are now stored as a “DocumentTermMatrix.” In this format, the matrix records the term (as equivalent of a column) and the document (as equivalent of row), and the number of times the term appears in the given document. Many terms will not appear in the document, meaning that the matrix will be stored as “sparse,” meaning there will be a preponderance of zeroes. Here, since we are looking only at two documents that both come from a single volume set, the sparsity is relatively low (only 27%). In most applications, the sparsity will be a lot higher, approaching 99% or more.

# (typing my analysis of the Alice model here)
#Here the sparsity is somewhat higher at 34% but is still low because we are only looking at two documents

#Estimating our topic model is then relatively simple. All we need to do if specify how many topics that we want to search for, and we can also set our seed, which is needed to reproduce the same results each time (as the model is a generative probabilistic one, meaning different random iterations will produce different results).

tocq_lda_alice <- LDA(tocq_dtm_alice, k = 5, control = list(seed = 1234))

tocq_topics_alice <-tidy(tocq_lda_alice, matrix= "beta")

#using only 5 topics here becaause Alice in Wonderland is more of a fiction book rather than a nonfiction book, and is also a lot shorter than de Tocqueville's book

head(tocq_topics_alice, n=5)


```

```{r}
#Alright so the betas above represent the probability that the given term belongs to a given topic. So, here, we see that the term “alice” most likely belongs to "topic 1." 
#We can then plots the top terms, in terms of beta, for each topic as follows:

tocq_top_terms_alice <- tocq_topics_alice %>%
  group_by(topic) %>%
  top_n(5, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)

tocq_top_terms_alice %>%
  mutate(term = reorder_within(term, beta, topic)) %>%
  ggplot(aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free", ncol = 4) +
  scale_y_reordered() +
  theme_tufte(base_family = "Helvetica")

#But we have to further evaluate these topics as here they all seem pretty similar 


```

```{r}

#Evaluating the topic model

#First we will plot the relative word frequencies and have a look at whether there are words that clearly distinguish the two volumes

tidy_tocq_alice <- tocq_alice %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words)

#count most common words in both

tidy_tocq_alice %>%
  count(word, sort= TRUE)

bookfreq_alice <- tidy_tocq_alice %>%
  mutate(booknumber = ifelse(gutenberg_id==11, "Alice_1", "Alice_2")) %>%
  mutate(word = str_extract(word, "[a-z']+")) %>%
  count(booknumber, word) %>%
  group_by(booknumber) %>%
  mutate(proportion = n / sum(n)) %>% 
  select(-n) %>% 
  spread(booknumber, proportion)

ggplot(bookfreq_alice, aes(x = Alice_1, y = Alice_2, color = abs(Alice_1 - Alice_2))) +
  geom_abline(color = "gray40", lty = 2) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient(limits = c(0, 0.001), low = "darkslategray4", high = "gray75") +
  theme_tufte(base_family = "Helvetica") +
  theme(legend.position="none", 
        strip.background = element_blank(), 
        strip.text.x = element_blank()) +
  labs(x = "Alice_1", y = "Alice_2") +
  coord_equal()

```

Words that further away from the best fit diagonal line appear with greater frequency in one book versus the other. So we see that there are differences between the first book "Alice's Adventures in Wonderland" versus the second book "Through the Looking Glass.' We see that cats and baby cats are more common in "Alice's Adventures in Wonderland," and that horses and the color red are more common in "Through the Looking Glass."

```{r}


#Remove NA text
tocq_alice <- tocq_alice %>%
  filter(!is.na(text))

# Assign book numbers and chapters- Alice_1 is "Alice's Adventures in Wonderland, Alice_2 is "Through the Looking Glass"
tocq_chapter_alice <- tocq_alice %>%
  mutate(booknumber = ifelse(gutenberg_id == 11, "Alice_1", "Alice_2")) %>%
  group_by(booknumber) %>%
  mutate(chapter = cumsum(str_detect(text, regex("^chapter ", ignore_case = TRUE)))) %>%
  ungroup() %>%
  filter(chapter > 0) %>%
  mutate(document = paste0(booknumber, "_chap", chapter)) 

#Tokenize the dataset into words
tocq_chapter_word_alice <- tocq_chapter_alice %>%
  unnest_tokens(word, text)

#Remove the stop words and count words by document
tocq_word_counts_alice <- tocq_chapter_word_alice %>%
  anti_join(stop_words, by = "word") %>%
  count(document, chapter, booknumber, word, sort = TRUE)

# Create Document-Term Matrix for LDA
tocq_chapters_dtm_alice <- tocq_word_counts_alice %>%
  cast_dtm(document, word, n)

# Inspect the sparsity
tm::inspect(tocq_chapters_dtm_alice)

# Fit LDA model with k=5 topics
tocq_chapters_lda_alice <- LDA(tocq_chapters_dtm_alice, k = 2, control = list(seed = 1234))

# Extract per-document topic probabilities (gamma)
alice_tocq_chapters_gamma <- tidy(tocq_chapters_lda_alice, matrix = "gamma")

# Check results
head(alice_tocq_chapters_gamma)


#Now that we have our topic probabilities, we can see how well our unsupervised learning did at distinguishing the two volumes generatively just from the words contained in each chapter

alice_tocq_chapters_gamma <- alice_tocq_chapters_gamma %>%
  separate(document, c("title", "chapter"), sep = "_", convert = TRUE)


```

```{r}
tocq_chapter_classifications_alice <- alice_tocq_chapters_gamma%>%
group_by(title, chapter)%>%
  top_n(1, gamma)%>%
  ungroup()
```

```{r}
tocq_book_topics_alice <- tocq_chapter_classifications_alice %>%
  count(title, topic) %>%
  group_by(title) %>%
  top_n(1, n) %>%
  ungroup() %>%
  transmute(consensus = title, topic)
```

```{r}

# Look document-word pairs were to see which words in each documents were assigned
# to a given topic


assignments_alice_km1 <- augment(tocq_chapters_lda_alice, data = tocq_chapters_dtm_alice)
head(assignments_alice)


```

```{r}
assignments_alice_km1 <- assignments_alice_km1 %>%
 separate(document, c("title", "chapter"), sep = "_", convert = TRUE) %>%
inner_join(tocq_book_topics_alice, by = c(".topic" = "topic"))


assignments_alice_km %>%
  count(title, .topic, wt = count) %>%
  complete(title, .topic, fill = list(n = 0)) %>%  # fills in zeros
  group_by(title) %>%
  mutate(percent = n / sum(n)) %>%
  ggplot(aes(.topic, title, fill = percent)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  geom_text(aes(label = scales::percent(percent))) +
  theme_minimal()+
  labs(x="Book words assigned to",
       y= "Book words came from",
       fill= "% of assignments")
```

Since more than half (56%) of Alice_1's words are assigned to topic 2, topic 2 likely represents content most strongly associated with Alice_1 The 44% assigned to topic 1 suggests that there is shared overlap and vocabulary between the two topics.This model behavior therefore aligns with the gamma values as it confirms that the chapters strongly associate with one topic though some vocabulary does cross the topic boundaries

Lastly we will do our validations

```{r}
# load in corpus of Alice in Wonderland/ Through the Looking Glass text data.
corp_alice <- corpus(tocq_alice, text_field = "text")
# use first 10 documents for example
documents_alice <- corp[sample(1:30000,1000)]
# take a look at the document names
print(names(documents_alice[1:10]))
```

```{r}
# load in corpus of Alice in Wonderland/ Through the Looking Glass text data.
corp_alice <- corpus(tocq_alice, text_field = "text")
# use first 10 documents for example
documents_alice <- corp[sample(1:30000,1000)]
# take a look at the document names
print(names(documents_alice[1:10]))
```

```{r}
#Now we will pre process the text

preprocessed_documents_alice <- factorial_preprocessing(
    documents_alice,
    use_ngrams = TRUE,
    infrequent_term_threshold = 0.2,
    verbose = FALSE)
```

We can then get the results of our pre-processing, comparing the distance between documents that have been processed in different ways.

```{r}
preText_results_alice <- preText(
    preprocessed_documents_alice,
    dataset_name = "Alice",
    distance_method = "cosine",
    num_comparisons = 20,
    verbose = FALSE)
```

```{r}
preText_results_alice<-preText(
  preprocessed_documents_alice,
  dataset_name = "Alice",
  distance_method = "cosine",
  num_comparisons = 20,
  verbose = FALSE
)

#and we can plot these accordingly 
preText_score_plot(preText_results_alice)

```
